---
title: "Understanding Data Science Jobs in Different Industries"
output: github_document
---

```{r, include = FALSE}
library(tidyverse)
library(readr)
library(viridis)
library(gt)
library(ggplot2)
library(viridis)
library(ggalluvial)
library(ggfittext)

ds_jobs <- read.csv("../2_Data/ds_jobs.csv")
```

In this file we are looking at the industry in which each job opening is located to understand if certain industries are represented more in certain areas. In addition, we are interested in seeing if there are certain industries that are only looking for certain data science roles or if all roles are needed within all industries in our data.

## Top Industries in terms of Data Science Job Postings

```{r, message = FALSE}
#filter for data required for graph
data_jobs <- ds_jobs %>% 
  group_by(job_category) %>% 
  summarize(count = n()) %>% 
  filter(!job_category %in% 
           c('Biologist', 'Consultant', 'Research Scientist', 
           'Computer Scientist', NA))

# find the counts of data science job postings within each industry
data_industries <- ds_jobs %>%
  filter(job_category %in% data_jobs$job_category) %>% 
  group_by(industry) %>% 
  summarize(count = n()) %>% 
  filter(!is.na(industry)) %>% 
  arrange(-count) %>% 
  filter(count >= 100)

data_industries
```

The top industries in terms of data science job postings are listed above. Information Technology and Business Services are far ahead of the rest of the top 5 in terms of the number of data science roles required: Biotech/Pharmaceuticals, Finance, Aerospace/Defense.

## Industry vs Location

We want to know which industries have data science jobs available in each metro area. We will stick to the top 5 industries in terms of the number of data science job openings for this analysis since there are too many industries in our data set to look at them all visually.

```{r, message  = FALSE, warning = FALSE}
ds_filter <- ds_jobs %>%
  filter(!is.na(job_category)) %>%
  filter(job_category == "Data Analyst" | job_category == "Data Engineer" | 
         job_category == "Data Scientist" | job_category == "Machine Learning Engineer" |
         job_category == "Statistician" | job_category == "Other Analyst")

ds_filter2 <- ds_filter %>%
  filter(!is.na(industry))
ds_filter3 <- ds_filter2 %>%
  group_by(industry) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  slice(1:5)%>%
  ungroup()
ds_filter2 <- ds_filter2 %>%
  filter(industry == ds_filter3$industry)

ggplot(ds_filter2, aes(metro_location,fill = industry)) +
  geom_bar() +
  scale_fill_viridis(discrete = TRUE)+
  labs(
    title = "Top 5 Industries by Location",
    x = "Location",
    subtitle = "The cities are listed with the total 
    number of jobs and it is colored by the industry"
  )+
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8),
    axis.text = element_text(size =10),
    plot.subtitle=element_text(size=8, color = "grey"),
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8)
  )
```

Filtering by the 6 data science categories from `job_category`, we can see that San Francisco and DC hold the majority of data science jobs. Now we can also see in which industries these data science jobs are so applicants can determine good industries for data scientists. Judging by this graph, IT seems to be where most of the jobs in these areas are being posted. This makes sense because many IT companies have a lot more data and require more roles within the data science field.

## Data Science Roles Breakdown by Industry

This alluvial plot gives us more of a visual explanation/view about the top 5 industries from our data set and how they map to the different data science jobs available. Where do most data scientists fit in? We see clearly that there are data scientists in every industry with the most in Business services. Although present in various industries, Machine learning and Statistics are narrowed down to a particular domain and the numbers are pretty low compared to data engineer or data analyst.

```{R, message = FALSE, warning = FALSE}
ds_filter <- ds_jobs %>%
  filter(!is.na(job_category)) %>%
  filter(job_category == "Data Analyst" | job_category == "Data Engineer" | job_category == "Data Scientist" | job_category == "Machine Learning Engineer" | job_category == "Statistician" | job_category == "Other Analyst")
ds_filter2 <- ds_filter %>%
  filter(!is.na(industry))
ds_filter3 <- ds_filter2 %>%
  group_by(industry) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  slice(1:5)%>%
  ungroup()
ds_filter2 <- ds_filter2 %>%
  filter(industry == ds_filter3$industry) %>%
  filter(!is.na(job_category)) %>%
  filter(job_category == "Data Analyst" | job_category == "Data Engineer" | job_category == "Data Scientist" | job_category == "Machine Learning Engineer" | job_category == "Statistician")

df <- table(ds_filter2$job_category, ds_filter2$industry)
dff <- data.frame(df)

ggplot(data = dff,
       aes(axis1 = Var1, axis2 = Var2,
           y = Freq)) +
  scale_x_discrete(limits = c("Job Category", "Industry"), expand = c(.2, .05)) +
  ylab("Frequency") +
  geom_alluvium(aes(fill = Var1), alpha = 0.9) +
  geom_stratum() +
  scale_fill_viridis(discrete = TRUE, name = "Data Science Role") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2) +
  theme_minimal() +
  ggtitle("Data Science Job Roles in the Top 5 Industries")
```
