---
title: "Next Steps"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)
```

# Read in Data

```{r}
NY <- read.csv("glassdoor_data/Data_Job_NY.csv")
SF <- read.csv("glassdoor_data/Data_Job_SF.csv")
TX <- read.csv("glassdoor_data/Data_Job_TX.csv")
WA <- read.csv("glassdoor_data/Data_Job_WA.csv")

NY["Region"] <- "NYC"
SF["Region"] <- "SF"
TX["Region"] <- "TX"
WA["Region"] <- "DC"

#combining data from the 4 regions
ds_jobs <- rbind(NY, SF, TX, WA)
```

# Basic Exploration

```{r}
dim(ds_jobs)
```
There are 3,324 rows and 12 variables in the data set.

```{r}
head(ds_jobs)
```
```{r}
names(ds_jobs)
```

```{r}
summary(ds_jobs)
```

The unit of analysis in the `ds_jobs` data set is a specific job within the field of data science at a company listed on glassdoor.

# Clean Data

## Missing Values

It looks like `Min_salary` and `Max_Salary` have missing values denoted with a `-1` value. We will replace them with NA. 

```{r}
#Replace -1 with NA
ds_jobs["Min_Salary"] <- na_if(ds_jobs["Min_Salary"], -1)
ds_jobs["Max_Salary"] <- na_if(ds_jobs["Max_Salary"], -1)
```

```{r}
table(ds_jobs$State)
```

There are 2 values within `State` that are blank. We will replace them with NA.

```{r}
#Replace blanks with NA
ds_jobs["State"] <- na_if(ds_jobs["State"], "")
```

```{r}
table(ds_jobs$Industry)
```

There are 624 values with no industry. We will replace them with NA.

```{r}
#Replace blanks with NA
ds_jobs["Industry"] <- na_if(ds_jobs["Industry"], "")
```

## Data Types

```{r}
sapply(ds_jobs, class)
```

It looks like all the data types are correct except for `Date_Posted` and `Valid_until` which should be dates.

```{r}
ds_jobs$Date_Posted <- as.Date(ds_jobs$Date_Posted, format = "%Y-%m-%d")  
ds_jobs$Valid_until <- as.Date(ds_jobs$Valid_until, format = "%Y-%m-%d")  
```



