---
title: "EDA"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

# Regions

We want to find all of the regions that are incorporated in this data set so we will group the city variable and calculate how many of each city there is.

```{r}
ds_jobs %>%
  group_by(City) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```

Having done this, we see that the larger cities are most often found. We want to incorporate the suburban areas around these big cities to get a better regional report of where the DS jobs are.

## NY data set
```{r}
NY %>%
  group_by(City) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```

Below is a list of the cities in the NY data set. Thus, the cities given in the data set are all surrounding cities to NYC.
NY metro area:
* New York
* Brooklyn
* Rego Park (Queens)
* Staten Island
* Bronx
* Fort Lee (in Jersey but would fit the mold of being a part of NYC)
* Jersey City (again, same as Fort Lee, just on opposite side of Hudson River)
* Lynbrook (on Long Island, 1 hour outside of Manhattan)
* Mamaroneck (~1 hour away from Manahattan)
* Maspeth (Queens)
* Paramus (in NJ just across Hudson)
* West New York
* West Orange (NJ, ~19 mi away from Manhattan)
* Williston Park (~22 mi away from Manhattan)

## TX data set

I want to know the count of each city so we know which parts of TX we are working with.

```{r}
TX %>%
  group_by(City) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

TX %>%
  filter(City == "")

ds_jobs["City"] <- na_if(ds_jobs["City"], "")
```

***NA value alert***
6 'City' values are blank, so we need to determine how we want to handle this. Currently, I have made them NA. I think we should remove them because they do not provide any value on where the jobs are located.

Because of the large number of cities used, do we want to use cities that only have 1 job?

```{r}
TX %>%
  group_by(City) %>%
  summarize(count = n()) %>%
  filter(count == 1)
```

* Bryan (next to College Station)
* Burnet (~50 mi outside of Austin)
* Chennai (I can't find it on the map)
* Comanche Village I (~70 mi outside Austin)
