---
title: "R Notebook"
output: github_document
---

# Load Packages and Data

```{r}
library(tidyverse)
library(ggplot2)
library(viridis)
```

```{r}
setwd('C:/Users/kingl/Desktop/Projects/eda_fall20/final_project/eda20-team4-project/')
ds_jobs <- read.csv('Data Cleaning/ds_jobs.csv')
```

# Top Companies in terms of Overall Job Postings

```{r}
companies <- ds_jobs %>%
  filter(is.na(job_category) == F) %>% 
  group_by(company) %>% 
  summarize(count = n()) %>% 
  arrange(-count) %>% 
  filter(count >= 30)

companies
```

```{r}
ds_jobs %>% 
  filter(is.na(job_category) == F,
         company %in% companies$company) %>% 
  group_by(company, job_category) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = company, y = count)) + geom_col(aes(fill = job_category), position = 'dodge2') +
  scale_fill_viridis(discrete = TRUE)
```

The above plot displays the job category count of the top 6 overall companies in terms of job postings. From here, we see that we may need to subset our data to only include data science specific jobs. 

# List Counts of Data Science Jobs

```{r}
data_jobs <- ds_jobs %>% 
  group_by(job_category) %>% 
  summarize(count = n()) %>% 
  filter(job_category %in% 
           c('Data Engineer', 'Data Analyst', 'Data Scientist', 
             'Machine Learning Engineer', 'Statistician', 'Other Analyst'))

data_jobs
```

In our dataset, we see most data science jobs listed under the data scientist role, followed closely by data engineer and data analyst. There are also roles of machine learning engineer, statistician and other analyst - though to a much lesser extent.

# Companies Analysis

## Top Companies in Terms of Data Science Job Postings

```{r}
data_companies <- ds_jobs %>%
  filter(job_category %in% data_jobs$job_category) %>% 
  group_by(company, industry, rating) %>% 
  summarize(count = n()) %>% 
  arrange(-count) %>% 
  filter(count >= 20)

data_companies
```

This tibble displays the top companies, their ratings, and their associated industry in terms of data science job postings. While there are various industries represented, Business Services companies appear three times above the 20 job cutoff. 

## Job Category Breakdown of Top Data Science Companies

```{r}
ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category,
         company %in% data_companies$company) %>% 
  group_by(company, job_category) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = company, y = count)) + 
  geom_col(aes(fill = job_category)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = 'Job Category Breakdown of Top Data Science Companies',
       subtitle = 'Companies with at least 20 DS job postings',
       x = 'Company',
       y = 'Count') +
  scale_fill_viridis(discrete = TRUE, name = 'Job Category')
```

From this plot we see that certain companies only have postings for one type of job: Addepar-Data Engineer, Kingdom Associates/TEECOM-Machine Learning Engineer, National Debt Relief-Data Analyst. Genetech not only has the most jobs available of the top companies, but they also have every kind of data science role available. CyberCoders, Booz Allen Hamilton Inc., and Management Decisions Inc. also have a variety of roles available. 

## Metro Area Breakdown of Top Data Science Companies

```{r}
ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category,
         company %in% data_companies$company) %>% 
  group_by(company, metro_location) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = company, y = count)) + 
  geom_col(aes(fill = metro_location)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = 'Metro Area Breakdown of Top Data Science Companies',
       subtitle = 'Companies with at least 20 DS job postings',
       x = 'Company',
       y = 'Count') +
  scale_fill_viridis(discrete = TRUE, name = 'Metro Area')
```

When examining the location of the job postings for the top data science companies, we see that most of the job postings for a single company are located in one area. Addeper, Kingdom Associates, National Debt Relief, and TEECOM are all located in NYC. Genetech is located exclusively in San Francisco. Booz Allen and Management Decisions are primarily located in one area, though they have a single digit number of postings in other locations. CyberCoders has job postings located in a variety of Metro Areas, matching their data science role versatility with their location versatility. 

# Industry Analysis

## Most Popular Industries in terms of Data Science Job Postings

```{r}
data_industries <- ds_jobs %>%
  filter(job_category %in% data_jobs$job_category) %>% 
  group_by(industry) %>% 
  summarize(count = n()) %>% 
  filter(!is.na(industry)) %>% 
  arrange(-count) %>% 
  filter(count >= 100)

data_industries
```

The top industries in terms of data science job postings are listed above. Information technology and business services are far ahead of the rest of the top 5: Biotech/Pharmaceuticals, Finance, Aerospace/Defense.

# Job Category Analysis

## Metro Area Breakdown of Job Categories

```{r}
ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category) %>% 
  group_by(metro_location) %>% 
  summarize(count = n())
```

```{r}
ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category) %>% 
  group_by(metro_location, job_category) %>% 
  summarize(count = n()) %>% 
  mutate(pct = count/sum(count) * 100) %>% 
  ggplot(aes(x = metro_location, y = pct)) + 
  geom_col(aes(fill = job_category), position = 'dodge2') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = 'Metro Area Breakdown of Data Science Job Categories',
       x = 'Metro Area',
       y = 'Percent of Jobs') +
  scale_fill_viridis(discrete = TRUE, name = 'Job Category')
```

```{r}
ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category) %>% 
  group_by(metro_location, job_category) %>% 
  summarize(count = n()) %>% 
  mutate(pct = count/sum(count) * 100) %>% 
  ggplot(aes(x = metro_location, y = count)) + 
  geom_col(aes(fill = job_category)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = 'Metro Area Breakdown of Data Science Job Categories',
       x = 'Metro Area',
       y = 'Count') +
  scale_fill_viridis(discrete = TRUE, name = 'Job Category')
```

The above two plots display the same information, just in different ways. San Francisco and Washington DC clearly have the highest amount of DS job postings. Dallas and Austin surprisingly have a higher amount of job postings than New York. Data analyst is the most popular DS job in Austin and San Francisco. Data Engineer is the most popular job posting in Dallas. Data Scientist is the most popular job posting in both San Antonio and DC. Machine Learning Engineer makes up exactly half of the postings in New York, but in terms of DS roles the city offers the least versatility - only offering Analyst, Engineer, and ML Engineer positions (perhaps mention people leaving area during covid). Statistician is in the minority of available jobs at each location aside from New York.

# Rating Analysis

## Rating by Metro Location

```{r}
ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category) %>% 
  select(company, metro_location, industry, rating) %>% 
  unique() %>% 
  ggplot(aes(x = metro_location, y = rating)) + 
  geom_violin()
```

The above violin plot shows the distribution of company ratings across metro area. While most areas look the same, New York has all ratings concentrated between 3.5 and ~4.25. This may signal that applicants can be pretty certain about what life at the company is like in New York as opposed to other metro areas with greater variability in company ratings. 

```{r}
ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category) %>% 
  select(company, metro_location, industry, rating) %>% 
  unique() %>% 
  ggplot(aes(x = rating)) + 
  geom_density(aes(fill = metro_location), alpha = .5) +
  labs(title = 'Glassdoor Rating Density Plot by Metro Area',
       x = 'Rating',
       y = 'Density') +
  scale_fill_viridis(discrete = TRUE, name = 'Metro Area')
```

This density plot presents the same information as the above violin plot. We see a high concentration just above 4 for New York.

## Rating by Job Category

```{r}
ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category) %>% 
  select(company, metro_location, industry, job_category, rating) %>% 
  unique() %>% 
  ggplot(aes(x = rating)) + 
  geom_density(aes(fill = job_category), alpha = .5) +
  labs(title = 'Glassdoor Rating Density Plot by Job Category',
       x = 'Rating',
       y = 'Density') +
  scale_fill_viridis(discrete = TRUE, name = 'Job Category')

ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category) %>% 
  select(company, metro_location, industry, job_category, rating) %>% 
  unique() %>% 
  ggplot(aes(x = job_category, y = rating)) + 
  geom_violin()
```

The above density and violin plots display the ratings of each company across job category. There does not seem to be much difference in ratings across job categories, however we do see an absence of negative outliers for the other analyst and statistician positions. This may not be a reliable approach because companies can appear across various roles and rating is specific to company.

## Rating by Industry

```{r}
ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category,
         industry %in% data_industries$industry) %>% 
  select(company, metro_location, industry, rating) %>% 
  unique() %>% 
  ggplot(aes(x = rating)) + 
  geom_density(aes(fill = industry), alpha = .5) +
  labs(title = 'Glassdoor Rating Density Plot by Industry',
       x = 'Rating',
       y = 'Density') +
  scale_fill_viridis(discrete = TRUE, name = 'Job Category')

ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category,
         industry %in% data_industries$industry) %>% 
  select(company, metro_location, industry, rating) %>% 
  unique() %>% 
  ggplot(aes(x = industry, y = rating)) + 
  geom_violin()
```

The above density and violin plots show the rating of each company across industry. The distributions all appear pretty similar with the exception of a high concentration just above 3.5 for Aerospace and Defense. Information Technology has the highest rating peak among the five largest industries, while Biotech and Pharmaceuticals has the highest number of 5 rated companies. 