---
title: "Job_Types"
output: github_document
---
```{r, message = FALSE}
library(tidyverse)
library(ggplot2)
library(viridis)
```
Job types :

Viewwing the job typs from the data set :

```{r}
ds_jobs <- read.csv("../Data Cleaning/ds_jobs.csv")
```

Below are the job types :

* Contractor
* Fulltime
* Intern
* Other
* Parttime
* Temporary

```{R}
data_jobs <- ds_jobs %>% 
  group_by(job_category) %>% 
  summarize(count = n()) %>% 
  filter(!job_category %in% c('Biology', 'Consultant', 'Research Scientist', 'Computer Scientist', NA))

ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category) %>% 
  group_by(job_type, metro_location) %>% 
  summarize(count = n()) %>% 
  mutate(pct = count/sum(count) * 100) %>% 
  ggplot(aes(x = metro_location, y = pct)) + 
  geom_col(aes(fill = job_type), position = 'dodge2') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = 'Location by type of data science jobs available',
       x = 'Location',
       y = 'Count') +
  scale_fill_discrete(name = 'Location')+
  theme_classic() +
  scale_fill_viridis(discrete = TRUE, name = "Job type")+
  theme( axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))

ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category) %>% 
  group_by(job_type, metro_location) %>% 
  summarize(count = n()) %>% 
  mutate(pct = count/sum(count) * 100) %>% 
  ggplot(aes(x = job_type, y = pct)) + 
  geom_col(aes(fill = metro_location), position = 'dodge2') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = 'Location by type of data science jobs available',
       x = 'Job Type',
       y = 'Count') +
  scale_fill_discrete(name = 'Location')+
  theme_classic() +
  scale_fill_viridis(discrete = TRUE, name = "Location")+
  theme( axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))
```


Viewing the number of job types in each state as a table

```{r}
table(ds_jobs$metro_location, ds_jobs$job_type)
```




# Relationship between states and types of jobs available

```{r}
ggplot(ds_jobs, aes(x=metro_location, y=job_type))+
  geom_jitter(alpha = 0.7) + labs(title ="Relationship between states and types of jobs available",
  y = "Type Of Job", 
  x = "state")+
theme_classic()
```


From the graph, we can see the distribution of the types of jobs availbale in each state. We see that, KY, NC and TN doesn't have more full time jobs when compared to other states and there are only couple of states which offer temporary positions. 


Further, we can explore about the sub categories in full time jobs, by taking data science related jobs from the other rmd file.

# Analysing the number of jobs in every category by taking the ds_jobs_bucket.So, here we have five categories Data Analyst, Data Engineer, Data Scientist, Machine Learning and Statistics. Finding out how many of these fall into different industries. In our data frame, we have 23 industries for which we can map these job categories.

```{R}
df <- data.frame(table(ds_jobs$industry,ds_jobs$job_category,ds_jobs$job_type))

p <- ggplot(data=df, aes(x=Var2, y=Freq, fill=Var1)) + geom_bar(stat="identity") +
  labs(title = "Job category vs Industry",
       x = "Job Category",
       y = "Count")

p + scale_fill_discrete(name="Industry") +theme(
  plot.title = element_text(color="black", size=14, face="bold",hjust = 0.5)
)
```




# Time series analysis

There are two columns Date_Posted and Valid_Until which denotes the time frame until which the job applications are open and denoting the recruitment time frame for candidates. On trying to plot the availabiloty of new jobs while considering the date posted, we find that it doesn't increase exponentially neither does it hits the bottom but it reaches it's peak after mid May denoting the max job availability.


```{R}
table_date_posted <- data.frame(table(ds_jobs$date_posted))
table_date_posted$Var1 <- as.Date(table_date_posted$Var1, format = "%Y-%m-%d") 
class(table_date_posted$Var1)
ggplot(table_date_posted, aes(x = Var1, y = Freq)) +
  geom_line() +
  labs(title = "Job availability vs Date posted",
       x = "Date",
       y = "Number of jobs")

```


# Salary by location


```{R}
library(viridis)
df <- ds_jobs %>%
             group_by(metro_location) %>%
             summarise(min = min(min_scaled_salary, na.rm = TRUE),max = max(max_scaled_salary, na.rm = TRUE))

salary_data_loc <- df %>% pivot_longer(
                cols = c(min, max),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE
)


p <-ggplot(salary_data_loc, aes(x = reorder(metro_location, -salary), y = salary, fill = type)) 
  p + geom_bar(stat = "identity", position = 'dodge')+
    labs(title = "Salary by location",
       x = "Location",
       y = "Scaled Salary") +
    scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
    scale_y_continuous(
    breaks = seq(0,1100000,50000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
    theme_classic() +
  theme( axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))
```


# Salary by job type