---
title: "Job_Types"
output: github_document
---
```{r, message = FALSE}
library(tidyverse)
library(ggplot2)
library(viridis)
library(ggalluvial)
```
Job types :

Viewwing the job typs from the data set :

```{r}
ds_jobs <- read.csv("../Data Cleaning/ds_jobs.csv")
```

Below are the job types :

* Contractor
* Fulltime
* Intern
* Other
* Parttime
* Temporary

```{R}
data_jobs <- ds_jobs %>% 
  group_by(job_category) %>% 
  summarize(count = n()) %>% 
  filter(!job_category %in% c('Biology', 'Consultant', 'Research Scientist', 'Computer Scientist', NA))

ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category) %>% 
  group_by(job_type, metro_location) %>% 
  summarize(count = n()) %>% 
  mutate(pct = count/sum(count) * 100) %>% 
  ggplot(aes(x = metro_location, y = pct)) + 
  geom_col(aes(fill = job_type), position = 'dodge2') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = 'Location by type of data science jobs available',
       x = 'Location',
       y = 'Count') +
  scale_fill_discrete(name = 'Location')+
  theme_classic() +
  scale_fill_viridis(discrete = TRUE, name = "Job type")+
  theme( axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))

ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category) %>% 
  group_by(job_type, metro_location) %>% 
  summarize(count = n()) %>% 
  mutate(pct = count/sum(count) * 100) %>% 
  ggplot(aes(x = job_type, y = pct)) + 
  geom_col(aes(fill = metro_location), position = 'dodge2') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = 'Location by type of data science jobs available',
       x = 'Job Type',
       y = 'Count') +
  scale_fill_discrete(name = 'Location')+
  theme_classic() +
  scale_fill_viridis(discrete = TRUE, name = "Location")+
  theme( axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))
```
Table interpretation of above visualization plots. From the data set, considering only the full time data science realted job availabilities, we see that San Francisco and Washington have the maximum number of jobs available while San Antonio has the least jobs comparitively.If we also look at the job category by location, we find similarities. Hence, if a person is looking for a job on data science field, then based on this data, it is suggested to for jobs in the bay area and DC.


Viewing the number of job types in each state as a table

```{r}
table(ds_jobs$metro_location, ds_jobs$job_type)
```

# Location by job type

```{R}
data_jobs <- ds_jobs %>% 
  group_by(job_category) %>% 
  summarize(count = n()) %>% 
  filter(!job_category %in% c('Biology', 'Consultant', 'Research Scientist', 'Computer Scientist', NA))

ds_jobs %>% 
  filter(job_category %in% data_jobs$job_category) %>% 
  group_by(job_type, metro_location) %>% 
  summarize(count = n()) %>% 
  mutate(pct = count/sum(count) * 100) %>% 
  ggplot(aes(x = metro_location, y = pct)) + 
  geom_col(aes(fill = job_type), position = 'dodge2') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = 'Location by type of data science jobs available',
       x = 'Location',
       y = 'Count') +
  scale_fill_discrete(name = 'Location')+
  theme_classic() +
  scale_fill_viridis(discrete = TRUE, name = "Job type")+
  theme( axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))

```

# Industry by job category 

This alluvial plot gives us more of a visual explanation/view about the top 5 industries from our data set and how it's mapped to the different data science jobs available. Where does most data scientists fit in? We see clearly that there are data scientists in every industry and the maximum in Business services. Since Machine learning and Statistics are narrowed down to a particular domain, although these are present in various industries, the numbers are pretty low compared to data engineer or data analyst. 

```{R}
ds_filter <- ds_jobs %>%
  filter(!is.na(job_category)) %>%
  filter(job_category == "Data Analyst" | job_category == "Data Engineer" | job_category == "Data Scientist" | job_category == "Machine Learning" | job_category == "Statistics" | job_category == "Other Analyst")
ds_filter2 <- ds_filter %>%
  filter(!is.na(industry))
ds_filter3 <- ds_filter2 %>%
  group_by(industry) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  slice(1:5)%>%
  ungroup()
ds_filter2 <- ds_filter2 %>%
  filter(industry == ds_filter3$industry) %>%
  filter(!is.na(job_category)) %>%
  filter(job_category == "Data Analyst" | job_category == "Data Engineer" | job_category == "Data Scientist" | job_category == "Machine Learning" | job_category == "Statistics" | job_category == "Other Analyst")

df <- table(ds_filter2$job_category, ds_filter2$industry)
dff <- data.frame(df)

ggplot(data = dff,
       aes(axis1 = Var1, axis2 = Var2,
           y = Freq)) +
  scale_x_discrete(limits = c("Job Category", "Industry"), expand = c(.2, .05)) +
  ylab("Frequency") +
  geom_alluvium(aes(fill = Var1)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
  ggtitle("Job category by industry")+
  scale_fill_discrete(name = "Job category")
```

# Time series analysis

This data about the job postings are listed from glassdoor scraped during covid time. According to the glassdoor findings, there were very few jobs posted around April end and there's a sudden rise the upcoming week. It's very clear that the job market was not stable during the pandemic.

```{R}
table_date_posted <- data.frame(table(ds_jobs$date_posted))
table_date_posted$Var1 <- as.Date(table_date_posted$Var1, format = "%Y-%m-%d") 
class(table_date_posted$Var1)
ggplot(table_date_posted, aes(x = Var1, y = Freq)) +
  geom_line() +
  labs(title = "Job availability vs Date posted",
       x = "Date",
       y = "Number of jobs")

```

# Salary by location


```{R}
library(viridis)
df <- ds_jobs %>%
             group_by(metro_location) %>%
             summarise(min = min(min_scaled_salary, na.rm = TRUE),max = max(max_scaled_salary, na.rm = TRUE))

salary_data_loc <- df %>% pivot_longer(
                cols = c(min, max),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE
)


p <-ggplot(salary_data_loc, aes(x = reorder(metro_location, -salary), y = salary, fill = type)) 
  p + geom_bar(stat = "identity", position = 'dodge')+
    labs(title = "Salary by location",
          subtitle = "Salary for all job posting in various locations.",
       x = "Location",
       y = "Scaled Salary") +
    scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
    scale_y_continuous(
    breaks = seq(0,1100000,50000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
    theme_classic() +
  theme( axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))
```


```{R}
ds_filter <- ds_jobs %>%
  filter(!is.na(job_category)) %>%
  filter(job_category == "Data Analyst" | job_category == "Data Engineer" | job_category == "Data Scientist" | job_category == "Machine Learning" | job_category == "Statistics" | job_category == "Other Analyst")

library(viridis)
df <- ds_filter %>%
             group_by(metro_location) %>%
             summarise(min = min(min_scaled_salary, na.rm = TRUE),max = max(max_scaled_salary, na.rm = TRUE))

salary_data_loc <- df %>% pivot_longer(
                cols = c(min, max),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE
)


p <-ggplot(salary_data_loc, aes(x = reorder(metro_location, -salary), y = salary, fill = type)) 
  p + geom_bar(stat = "identity", position = 'dodge')+
    labs(title = "Salary by location",
         subtitle = "Salary for data science related job posting in various locations.",
       x = "Location",
       y = "Scaled Salary") +
    scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
    scale_y_continuous(
    breaks = seq(0,1100000,50000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
    theme_classic() +
  theme( axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))



```
# Salary by job type - Data science related jobs

```{r}
scaled_salary_data <- ds_jobs %>% select(min_scaled_salary, max_scaled_salary, job_category, industry, job_type) %>% pivot_longer(
                cols = c(min_scaled_salary, max_scaled_salary),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE
)

scaled_salary_data <- scaled_salary_data %>% mutate(type = case_when(
  type == "min_scaled_salary" ~ "min",
  type == "max_scaled_salary" ~ "max"
))

scaled_salary_data_jc <- scaled_salary_data %>% filter(!is.na(job_type))
ds_related <- c("Data Scientist", "Data Analyst", "Data Engineer", "Machine Learning Engineer", "Statistician", "Other Analyst")

scaled_salary_data_jc_related <- scaled_salary_data_jc %>% 
  filter(job_category %in% ds_related)

mean_salary_jc <- scaled_salary_data_jc_related %>% group_by(type, job_type) %>% 
  mutate(mean_rate = mean(salary))

scaled_salary_data_jc_related %>% ggplot(aes(x = salary, fill = type)) +
  geom_density(alpha = 0.60) +
  labs(
    title = "Salary by Job Type",
    x = "Scaled Salary",
    y = "Density") +
  scale_x_continuous(
    breaks = seq(15000,260000,75000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
   geom_vline(aes(xintercept=mean_salary_jc$mean_rate, col = type, group = job_type), linetype = "dashed", show.legend = FALSE) +
scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
scale_color_viridis(discrete = TRUE, begin = 0.25, end = 0.5, guide = FALSE) +
  geom_text(data = mean_salary_jc, aes(x = ifelse(type == "max", mean_rate+25000, mean_rate-25000), y = 2.4e-05, label = paste0('$', round(mean_rate/1000,1), 'K'), group = job_type, col = type), size = 2) +
  facet_wrap(~job_type) + 
  theme_classic() +
  theme(strip.text = element_text(face = "bold", size = 7),
        axis.text.x = element_text(size = 6))
```

