---
title: "Salary Based on Job Category, Industry and Metro Area"
output: github_document
---

```{r, include = FALSE}
library(tidyverse)
library(ggplot2)
library(viridis)

ds_jobs <- read.csv("../Data Cleaning/ds_jobs.csv")
```

# Metro Area

```{r}
ds_filter <- ds_jobs %>%
  filter(!is.na(job_category)) %>%
  filter(job_category == "Data Analyst" | job_category == "Data Engineer" | job_category == "Data Scientist" | job_category == "Machine Learning Engineer" | job_category == "Statistician" | job_category == "Other Analyst")

df <- ds_filter %>%
             group_by(metro_location) %>%
             summarise(min = mean(min_salary, na.rm = TRUE),
                       max = mean(max_salary, na.rm = TRUE))

salary_data_loc <- df %>% pivot_longer(
                cols = c(min, max),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE
)


p <- ggplot(salary_data_loc, aes(x = reorder(metro_location, -salary), y = salary, fill = type)) 
p + geom_bar(stat = "identity", position = 'dodge')+
    labs(title = "Average Salary for Data Science Jobs",
         subtitle = "Glassdoor Job Postings by Metro Area",
       x = "",
       y = "Average Salary") +
    scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
    scale_y_continuous(
    breaks = seq(0,1100000,50000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
  geom_text(aes(label=paste0('$', round(salary/1000,1), 'K')), position=position_dodge(width=0.9), vjust=-0.25, size = 2.5) +
    theme_classic() +
  theme( axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 9))

```


# Job Category

Is there a difference in salary between the different types of roles within data science? 

We want to look at different types of jobs within the data science field to see if there is a difference in salary between different roles. First we will look at the distribution of the minimum and maximum salary for each job category to understand their distributions. We used the scaled salary variables in this instance because we are looking at comparing salaries of jobs across all metro areas.

```{r}
#wrangling the data to fit both salaries on the same graph
scaled_salary_data <- ds_jobs %>% select(min_scaled_salary, max_scaled_salary, job_category, industry, job_type) %>% pivot_longer(
                cols = c(min_scaled_salary, max_scaled_salary),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE
)

#cleaning up the salary type
scaled_salary_data <- scaled_salary_data %>% mutate(type = case_when(
  type == "min_scaled_salary" ~ "min",
  type == "max_scaled_salary" ~ "max"
))

#create vector of jobs closely related to data science
ds_related <- c("Data Scientist", "Data Analyst", "Data Engineer", "Machine Learning Engineer", "Statistician", "Other Analyst")
```

```{r}
#remove NAs from job_category
scaled_salary_data_jc <- scaled_salary_data %>% filter(!is.na(job_category))

#density graph by job category
scaled_salary_data_jc %>% ggplot(aes(x = salary, fill = type)) +
  geom_density(alpha = 0.8) +
  labs(
    title = "Salary by Job Category",
    x = "Scaled Salary",
    y = "Density") +
  scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) +
  scale_x_continuous(
    breaks = seq(15000,400000,100000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
  facet_wrap(~job_category) + 
  theme_classic() +
  theme(strip.text = element_text(face = "bold", size = 7),
        axis.text.x = element_text(size = 6))
```

At first glance all the distributions seem similar. They are approximately normally distributed, but are all a bit right skewed. We can look at boxplots of the same data to see the summary statistics a bit better.

```{r}
#boxplots for each job category
scaled_salary_data_jc %>% ggplot(aes(y = salary, x = job_category, fill = type)) +
  geom_boxplot(alpha = 0.85) +
  labs(
    title = "Salary by Job Category",
    y = "Salary",
    x = "") +
  scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) +
  scale_y_continuous(
    breaks = seq(15000,400000,75000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
  theme_classic() +
   coord_flip()
```

We can see in most instances there are outliers on the upper end of the distribution. Even with the outliers, there does not seem to be too much difference in salary distribution for different job categories.

This data contains observations for several different job types such as full time, part time, contractor, etc. It does not make sense to compare the salary ranges of certain job types such as full time versus part time so we will look at only jobs that are typically a full 40 hours a week.

```{r}
#remove part time jobs 
scaled_salary_data_jc <- scaled_salary_data_jc %>% filter(job_type != "PART_TIME")
```

To get a better a understanding of a comparison of salaries between different roles within data science, we want to look at the average for maximum and minimum scaled salaries based on job roles.

```{r, message = FALSE}
#find average min and max salary for each job category
jc_avg_salary <- ds_jobs %>% filter(!is.na(job_category) & !is.na(min_scaled_salary) & !is.na(max_scaled_salary)) %>% 
  group_by(job_category) %>% 
  summarise(avg_max_salary = mean(max_scaled_salary), 
            avg_min_salary = mean(min_scaled_salary)) %>% ungroup()

#pivot data for ease of graphing
jc_avg_salary_long <- jc_avg_salary %>% pivot_longer(cols = c(avg_max_salary, avg_min_salary),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE)

#bar plot of salary by job category
jc_avg_salary_long %>% ggplot(aes(x = reorder(job_category, -salary), y = salary ,fill = type))+
  geom_bar(stat = "identity", position = 'dodge') +
  labs(
    title = "Average Salary by Job Category",
    x = "",
    y = "Scaled Salary") +
  scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
  scale_y_continuous(
    breaks = seq(0,110000,25000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) + 
  geom_text(aes(label=paste0('$', round(salary/1000,1), 'K')), position=position_dodge(width=0.9), vjust=-0.25, size = 2) +
  theme_classic() +
  theme( axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))
```

Here we can see that the highest paying role is Data Scientist based on both minimum and maximum salary and the lowest is Data Analyst based on the maximum salary and Computer Scientist based on the minimum of the salary range.

So far the analysis has been for all job roles, but since there are so many we wanted to narrow the analysis to just those job roles that are most closely related to data science. 

## Most Related to Data Science

The roles most related to the field of data science are Data Scientist, Data Analyst, Data Engineer, Machine Learning Engineer, Statistician and Other Analyst. We can look at the distribution of salary for only these roles.

```{r, message= FALSE}
#filter for jobs in ds related jobs vector
scaled_salary_data_jc_related <- scaled_salary_data_jc %>% 
  filter(job_category %in% ds_related)

#calculate mean salary by job category to use in graph
mean_salary_jc <- scaled_salary_data_jc_related %>% group_by(type, job_category) %>% 
  mutate(mean_rate = mean(salary))

#density graph for ds related jobs
scaled_salary_data_jc_related %>% ggplot(aes(x = salary, fill = type)) +
  geom_density(alpha = 0.60) +
  labs(
    title = "Distribution of Scaled Salary for Data Science Roles",
    x = "Scaled Salary",
    y = "Density") +
  scale_x_continuous(
    breaks = seq(15000,260000,75000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
   geom_vline(aes(xintercept=mean_salary_jc$mean_rate, col = type, group = job_category), linetype = "dashed", show.legend = FALSE) +
scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
scale_color_viridis(discrete = TRUE, begin = 0.25, end = 0.5, guide = FALSE) +
  geom_text(data = mean_salary_jc, aes(x = ifelse(type == "max", mean_rate+25000, mean_rate-25000), y = 2.4e-05, label = paste0('$', round(mean_rate/1000,1), 'K'), group = job_category, col = type), size = 2) +
  facet_wrap(~job_category) + 
  theme_classic() +
  theme(strip.text = element_text(face = "bold", size = 7),
        axis.text.x = element_text(size = 6))
```

The distribution of both minimum and maximum salary is pretty similar for all of the different jobs roles. In addition, the ratio or proportion of minimum salaries overlapping with maximum salaries is fairly similar between the different job categories.

```{r}
#filter for jobs in ds related jobs vector
jc_avg_salary_related <- jc_avg_salary_long %>% 
  filter(job_category %in% ds_related)

#bar plot of salary by job category for ds jobs
jc_avg_salary_related %>% ggplot(aes(x = reorder(job_category, -salary), y = salary ,fill = type)) +
  geom_bar(stat = "identity", position = 'dodge') +
  labs(
    title = "Salary for Data Science Roles",
    subtitle = "Glassdoor Job Openings",
    x = "",
    y = "Average Scaled Salary") +
scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
  scale_y_continuous(
    breaks = seq(0,110000,20000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
  geom_text(aes(label=paste0('$', round(salary/1000,1), 'K')), position=position_dodge(width=0.9), vjust=-0.25, size = 2.5) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))
```

In looking at the average minimum and maximum salary for each job category we can see a distinct pattern. The minimum and maximum have the same levels per job category with Data Scientist having the highest average maximum and minimum salary then Data Engineer and so on until Data Analyst which has the lowest average maximum and minimum salary compared to other data science related jobs. For those looking for the highest paying jobs in data science, they should look at Data Scientist and Data Engineer roles first. 

# Job Industry

Is there a salary difference based on the industry the data science job is in?

We will look at the top 5 industries and compare their scaled minimum and maximum salary ranges to see if there is a difference in salaries based on the type of industry a job is in. 

```{r}
#remove NAs from industry
scaled_salary_data_ji <- scaled_salary_data %>% filter(!is.na(industry))

#filter out part time jobs
scaled_salary_data_ji <- scaled_salary_data_ji %>% filter(job_type != "PART_TIME")

#find top 5 industries for ds jobs by count of ds jobs in each industry
scaled_salary_data_ji_related <- scaled_salary_data_ji %>% 
  group_by(industry) %>% 
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  slice(1:5) %>%
  select(industry) %>% 
  ungroup()

#filter for jobs in top 5 industries
scaled_salary_data_ji <- scaled_salary_data_ji %>% 
  filter(industry %in% scaled_salary_data_ji_related$industry)

#calculate mean salary by job category to use in graph
mean_salary_ji <- scaled_salary_data_ji %>% group_by(type, industry) %>% 
  mutate(mean_rate = mean(salary))

#density plot of salary for top 5 industries
scaled_salary_data_ji %>% ggplot(aes(x = salary, fill = type)) +
  geom_density(alpha = 0.6) +
  labs(
    title = "Distribution of Scaled Salary for Top 5 Industries",
    x = "Scaled Salary",
    y = "Density") +
  scale_x_continuous(
    breaks = seq(15000,400000,100000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
  geom_vline(aes(xintercept=mean_salary_ji$mean_rate, col = type, group = industry), linetype = "dashed", show.legend = FALSE) +
scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
  scale_color_viridis(discrete = TRUE, begin = 0.25, end = 0.5, guide = FALSE) +
  geom_text(data = mean_salary_ji, aes(x = ifelse(type == "max", mean_rate+35000, mean_rate-35000), y = 2.5e-05, label = paste0('$', round(mean_rate/1000,1), 'K'), group = industry, col = type), size = 2) +
  facet_wrap(~industry) + 
  theme_classic() +
  theme(strip.text = element_text(face = "bold", size = 7),
    axis.text.x = element_text(size = 6))
```

At first glance, it does not look like there is much of a difference in distribution of minimum or maximum salary ranges between different industries. Again just like with job categories it seems like the salarly distributions are approximately normally distributed with a right skew.

```{r}
#calculate average min and max salary for each industry
ji_avg_salary <- ds_jobs %>% filter(!is.na(industry) & !is.na(min_scaled_salary & job_type != "PART_TIME") & !is.na(max_scaled_salary)) %>% 
  group_by(industry) %>% 
  summarise(avg_max_salary = mean(max_scaled_salary), 
            avg_min_salary = mean(min_scaled_salary)) %>% ungroup()

#pivot data for ease of graphing
ji_avg_salary_long <- ji_avg_salary %>% pivot_longer(cols = c(avg_max_salary, avg_min_salary),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE)

#filter for top 5 industries
ji_avg_salary_long_related <- ji_avg_salary_long %>% 
  filter(industry %in% scaled_salary_data_ji_related$industry)

#barplot of avg min and max salary by industry
ji_avg_salary_long_related %>% ggplot(aes(x = reorder(industry, -salary), y = salary ,fill = type)) +
  geom_bar(stat = "identity", position = 'dodge') +
  labs(
    title = "Average Salary for the Top 5 Industries",
    subtitle = "Data Science Related Job Openings",
    x = "",
    y = "Average Scaled Salary") +
scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
  scale_y_continuous(
    breaks = seq(0,100000,20000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
  geom_text(aes(label=paste0('$', round(salary/1000,1), 'K')), position=position_dodge(width=0.9), vjust=-0.25, size = 2.5) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))
```

There does seem to be a trend in the average minimum and maximum salaries for the top 5 industries. Information Technology has the highest average minimum salary followed by Finance with a slightly higher average maximum salary, then Business Services, Biotech & Pharmaceuticals and lastly Health Care. If an individual is looking for a high paying job in the data science field, they should look at these 5 industries that have the highest average salaries.

# Metro Area

The above analysis is for all full time data science jobs across all the metro locations in our data set. Are salaries for different roles in data science the same in different metro locations?

```{r, message = FALSE}
ds_jobs %>% filter(job_type != "PART_TIME" & job_category %in% ds_related) %>% group_by(metro_location) %>% 
  summarise(count = n()) %>% arrange(-count)
```

We can see that there are significantly more full time jobs in San Francisco, Washington DC and New York City. We want to look at salaries of different data science roles within each metro area to see if there are difference to the overall data set.

## San Francisco

We want to filter the data set for only full time data science related jobs in San Francisco and look at the distribution of minimum and maximum salary as well as average salaries across different data science related roles. For this analysis we are using the salary, not the scaled salary, as we are only look at each metro location separately and do not need to account for cost of living in the comparison.

```{r}
#filtering for SF data
sf_data <- ds_jobs %>% filter(job_type != "PART_TIME" & 
                              metro_location == "San Francisco, CA" &
                              job_category %in% ds_related) 

#wrangling the data to fit both salaries on the same graph
salary_data_sf <- sf_data %>% select(min_salary, max_salary, job_category, industry, job_type, metro_location) %>% 
  pivot_longer(
                cols = c(min_salary, max_salary),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE)

#calculate mean salary by job category to use in graph
mean_salary_sf <- salary_data_sf %>% group_by(type, job_category) %>% 
  mutate(mean_rate = mean(salary))

#density graph for ds related jobs
salary_data_sf %>% ggplot(aes(x = salary, fill = type)) +
  geom_density(alpha = 0.60) +
  labs(
    title = "Salary by Job Category",
    subtitle = "San Francisco Bay Metro Area",
    x = "Salary",
    y = "Density") +
  scale_x_continuous(
    breaks = seq(30000,320000,75000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
   geom_vline(aes(xintercept=mean_salary_sf$mean_rate, col = type, group = job_category), linetype = "dashed", show.legend = FALSE) +
scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
scale_color_viridis(discrete = TRUE, begin = 0.25, end = 0.5, guide = FALSE) +
  geom_text(data = mean_salary_sf, aes(x = ifelse(type == "max_salary", mean_rate+35000, mean_rate-35000), y = 1.8e-05, label = paste0('$', round(mean_rate/1000,1), 'K'), group = job_category, col = type), size = 2) +
  facet_wrap(~job_category) + 
  theme_classic() +
  theme(strip.text = element_text(face = "bold", size = 7),
        axis.text.x = element_text(size = 6))
```

The shape of the distribution for San Francisco looks similar to the distribution of salary for all metro areas although it looks less skewed to the right.

```{r, message = FALSE}
#calculating the average min and max salary for SF
sf_avg_salary <- sf_data %>% filter(!is.na(min_salary) & !is.na(max_salary)) %>% 
  group_by(job_category) %>% 
  summarise(avg_max_salary = mean(max_salary), 
            avg_min_salary = mean(min_salary))

#pivot data for ease of graphing
sf_avg_salary_long <- sf_avg_salary %>% pivot_longer(cols = c(avg_max_salary, avg_min_salary),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE)

#bar plot of salary by job category
sf_avg_salary_long %>% ggplot(aes(x = reorder(job_category, -salary), y = salary ,fill = type))+
  geom_bar(stat = "identity", position = 'dodge') +
  labs(
    title = "Average Salary for Data Science Roles",
    subtitle = "San Francisco Bay Metro Area",
    x = "",
    y = "Average Salary") +
  scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
  scale_y_continuous(
    breaks = seq(0,190000,40000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) + 
  geom_text(aes(label=paste0('$', round(salary/1000,1), 'K')), position=position_dodge(width=0.9), vjust=-0.25, size = 2) +
  theme_classic() +
  theme( axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))
```

In the San Francisco Bay area it looks like the role with the highest average minimum and maximum salary is a machine learning engineer. The average minimum is surprisingly high at $131.9K, but we do know that these salaries are a bit higher due to San Francisco having the highest cost of living.

```{r, message = FALSE}
sf_salary_industries <- salary_data_sf %>% filter(!is.na(industry))%>%                                                            group_by(industry) %>% 
                                           summarise(count = n()) %>%
                                           arrange(desc(count)) %>%
                                           slice(1:5) %>%
                                           select(industry) %>% 
                                           ungroup()

#filter for jobs in top 5 industries
sf_ji <- salary_data_sf %>% 
  filter(industry %in% sf_salary_industries$industry & !is.na(industry))

#calculate mean salary by job category to use in graph
mean_sf_ji <- sf_ji %>% group_by(type, industry) %>% 
  mutate(mean_rate = mean(salary))

#density plot of salary for top 5 industries
sf_ji %>% ggplot(aes(x = salary, fill = type)) +
  geom_density(alpha = 0.6) +
  labs(
    title = "Salary by Job Industry",
    subtitle = "San Francisco Bay Metro Area",
    x = "Salary",
    y = "Density") +
  scale_x_continuous(
    breaks = seq(20000,385000,100000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
  geom_vline(aes(xintercept=mean_sf_ji$mean_rate, col = type, group = industry), linetype = "dashed", show.legend = FALSE) +
scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
  scale_color_viridis(discrete = TRUE, begin = 0.25, end = 0.5, guide = FALSE) +
  geom_text(data = mean_sf_ji, aes(x = ifelse(type == "max_salary", mean_rate+40000, mean_rate-40000), y = 1.6e-05, label = paste0('$', round(mean_rate/1000,1), 'K'), group = industry, col = type), size = 2) +
  facet_wrap(~industry) + 
  theme_classic() +
  theme(strip.text = element_text(face = "bold", size = 7),
    axis.text.x = element_text(size = 6))
```

The top 5 industries in the San Francisco Bay area are the same as the top 5 industries across all metro areas in the data. 

```{r, message = FALSE}
#calculate average min and max salary for each industry
ji_sf_avg <- sf_data %>% filter(!is.na(min_salary) & !is.na(max_salary) &
                                industry %in% sf_salary_industries$industry) %>% 
                                group_by(industry) %>% 
                                summarise(avg_max_salary = mean(max_salary), 
                                avg_min_salary = mean(min_salary))

#pivot data for ease of graphing
ji_sf_avg_long <- ji_sf_avg %>% pivot_longer(cols = c(avg_max_salary, avg_min_salary),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE)

#barplot of avg min and max salary by industry
ji_sf_avg_long %>% ggplot(aes(x = reorder(industry, -salary), y = salary ,fill = type)) +
  geom_bar(stat = "identity", position = 'dodge') +
  labs(
    title = "Average Salary for the Top 5 Industries",
    subtitle = "San Francisco Bay Metro Area",
    x = "",
    y = "Average Salary") +
scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
  scale_y_continuous(
    breaks = seq(0,160000,25000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
  geom_text(aes(label=paste0('$', round(salary/1000,1), 'K')), position=position_dodge(width=0.9), vjust=-0.25, size = 2.5) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))
```
The industry with the highest average maximum and minimum salary in the San Francisco Bay area is Information Technology. The lowest average maximum salary is Health Care with the lowest average minimum salary in Business Services although these two industries have very close ranges in average salaries overall. 

## Washington DC

We want to filter the data set for only full time data science related jobs in Washington D.C.  and look at the distribution of minimum and maximum salary as well as average salaries across different data science related roles. For this analysis we are using the salary, not the scaled salary, as we are only look at each metro location separately and do not need to account for cost of living in the comparison.

```{r, message=FALSE}
#filtering for DC data
dc_data <- ds_jobs %>% filter(job_type != "PART_TIME" & 
                              metro_location == "Washington, DC" &
                              job_category %in% ds_related) 

dc_data %>% group_by(job_category) %>% 
  summarise(count = n())
```
Due to there only being one Other Analyst job role in Washington, D.C., we will remove this category from our analysis as it does not make sense to compare the density of different job categories for just one data point.

```{r}
dc_data <- dc_data %>% filter(job_category != "Other Analyst")

#wrangling the data to fit both salaries on the same graph
salary_data_dc <- dc_data %>% select(min_salary, max_salary, job_category, industry, job_type, metro_location) %>% 
  pivot_longer(
                cols = c(min_salary, max_salary),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE)

#calculate mean salary by job category to use in graph
mean_salary_dc <- salary_data_dc %>% group_by(type, job_category) %>% 
  mutate(mean_rate = mean(salary))

#density graph for ds related jobs
salary_data_dc %>% ggplot(aes(x = salary, fill = type)) +
  geom_density(alpha = 0.60) +
  labs(
    title = "Salary by Job Category",
    subtitle = "Washington D.C. Metro Area",
    x = "Salary",
    y = "Density") +
  scale_x_continuous(
    breaks = seq(20000,240000,75000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
   geom_vline(aes(xintercept=mean_salary_dc$mean_rate, col = type, group = job_category), linetype = "dashed", show.legend = FALSE) +
scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
scale_color_viridis(discrete = TRUE, begin = 0.25, end = 0.5, guide = FALSE) +
  geom_text(data = mean_salary_dc, aes(x = ifelse(type == "max_salary", mean_rate+35000, mean_rate-35000), y = 1.8e-05, label = paste0('$', round(mean_rate/1000,1), 'K'), group = job_category, col = type), size = 2) +
  facet_wrap(~job_category) + 
  theme_classic() +
  theme(strip.text = element_text(face = "bold", size = 7),
        axis.text.x = element_text(size = 6))
```

The shape of the distribution for Washington DC looks similar to the distribution of salary for all metro areas although it looks less skewed to the right.

```{r, message = FALSE}
#calculating the average min and max salary for DC
dc_avg_salary <- dc_data %>% filter(!is.na(min_salary) & !is.na(max_salary)) %>% 
  group_by(job_category) %>% 
  summarise(avg_max_salary = mean(max_salary), 
            avg_min_salary = mean(min_salary))

#pivot data for ease of graphing
dc_avg_salary_long <- dc_avg_salary %>% pivot_longer(cols = c(avg_max_salary, avg_min_salary),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE)

#bar plot of salary by job category
dc_avg_salary_long %>% ggplot(aes(x = reorder(job_category, -salary), y = salary ,fill = type))+
  geom_bar(stat = "identity", position = 'dodge') +
  labs(
    title = "Average Salary for Data Science Roles",
    subtitle = "Washington D.C. Metro Area",
    x = "",
    y = "Average Salary") +
  scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
  scale_y_continuous(
    breaks = seq(0,165000,40000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) + 
  geom_text(aes(label=paste0('$', round(salary/1000,1), 'K')), position=position_dodge(width=0.9), vjust=-0.25, size = 2) +
  theme_classic() +
  theme( axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))
```

Like the overall distribution, the data science related job with the highest average maximum and minimum salary in Washington D.C. is a Data Scientist.

```{r, message = FALSE}
dc_salary_industries <- salary_data_dc %>% filter(!is.na(industry))%>%                                                            group_by(industry) %>% 
                                           summarise(count = n()) %>%
                                           arrange(desc(count)) %>%
                                           slice(1:5) %>%
                                           select(industry) %>% 
                                           ungroup()

#filter for jobs in top 5 industries
dc_ji <- salary_data_dc %>% 
  filter(industry %in% dc_salary_industries$industry & !is.na(industry))

#calculate mean salary by job category to use in graph
mean_dc_ji <- dc_ji %>% group_by(type, industry) %>% 
  mutate(mean_rate = mean(salary))

#density plot of salary for top 5 industries
dc_ji %>% ggplot(aes(x = salary, fill = type)) +
  geom_density(alpha = 0.6) +
  labs(
    title = "Salary by Job Industry",
    subtitle = "Washington D.C. Metro Area",
    x = "Salary",
    y = "Density") +
  scale_x_continuous(
    breaks = seq(20000,240000,100000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
  geom_vline(aes(xintercept=mean_dc_ji$mean_rate, col = type, group = industry), linetype = "dashed", show.legend = FALSE) +
scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
  scale_color_viridis(discrete = TRUE, begin = 0.25, end = 0.5, guide = FALSE) +
  geom_text(data = mean_dc_ji, aes(x = ifelse(type == "max_salary", mean_rate+40000, mean_rate-40000), y = 1.6e-05, label = paste0('$', round(mean_rate/1000,1), 'K'), group = industry, col = type), size = 2) +
  facet_wrap(~industry) + 
  theme_classic() +
  theme(strip.text = element_text(face = "bold", size = 7),
    axis.text.x = element_text(size = 6))
```

The top 5 industries in the Washington D.C. metro area are Aerospace & Defense, Biotech & Pharmaceuticals, Business Services, Government and Information Technology. The spread of salary ranges are approximately distributed. This is different than the overall top 5 industries which makes sense because there is a much larger government presence in D.C. making it a leader in both the Government and Aerospace & Defense industries. 

```{r, message = FALSE}
#calculate average min and max salary for each industry
ji_dc_avg <- dc_data %>% filter(!is.na(min_salary) & !is.na(max_salary) &
                                industry %in% dc_salary_industries$industry) %>% 
                                group_by(industry) %>% 
                                summarise(avg_max_salary = mean(max_salary), 
                                avg_min_salary = mean(min_salary))

#pivot data for ease of graphing
ji_dc_avg_long <- ji_dc_avg %>% pivot_longer(cols = c(avg_max_salary, avg_min_salary),
                names_to = "type",
                values_to = "salary",
                values_drop_na = TRUE)

#barplot of avg min and max salary by industry
ji_dc_avg_long %>% ggplot(aes(x = reorder(industry, -salary), y = salary ,fill = type)) +
  geom_bar(stat = "identity", position = 'dodge') +
  labs(
    title = "Average Salary for the Top 5 Industries",
    subtitle = "Washington D.C. Metro Area",
    x = "",
    y = "Average Salary") +
scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.5, name = "Salary Type",
                     labels = c("Max", "Min")) + 
  scale_y_continuous(
    breaks = seq(0,130000,25000),
    labels = function(x){paste0('$', x/1000, 'K')}
  ) +
  geom_text(aes(label=paste0('$', round(salary/1000,1), 'K')), position=position_dodge(width=0.9), vjust=-0.25, size = 2.5) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.95, size = 8))
```

The industry with the highest average salaries in the Washington, D.C. metro area is Aerospace & Defense. 

## New York City

While it would be interesting to look at the salaries for different job categories and industries in New York City metro area, unfortunately we do not have enough data in this data set to do a proper analysis. This is a limitation as if we were able to access more data we might have been able to incorporate a more thorough and granular analysis of the data science job market in New York City. 